---
title: "Lab 03 - Nobel laureates"
author: "Shatavia Bellmon"
date: "1-31-26"
output: github_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
nobel <- read_csv("data/nobel.csv")
```

## Exercises

Here is a link to the [lab instructions](https://datascience4psych.github.io/DataScience4Psych/lab03.html).

### Exercise 1

view(nobel)

The nobel dataset contains `r nrow(nobel)` observations and `r ncol(nobel)` variables. Each row represents a a Nobel Prize winner and their information such as why they won the award and personal demographics.

### Exercise 2

```{r}

nobel_living <- nobel %>% 
  filter(
    is.na(died_date),
    !is.na(country),
    gender != "org"
  )
```

After filtering for living laureates with available country data and excluding organizations, the resulting data frame contains `r nrow(nobel_living)` observations.

### Exercise 3

```{r}
nobel_living <- nobel_living %>%
  mutate(
    country_us = if_else(country == "USA", "USA", "Other")
  )
```

```{r}
nobel_living_science <- nobel_living %>%
  filter(category %in% c("Physics", "Medicine", "Chemistry", "Economics"))
```

```{r}
ggplot(nobel_living_science, aes(x = country_us)) +
  geom_bar() +
  facet_wrap(~ category) +
  coord_flip()
```

In every category represented in this faceted bar plot, there are more nobel prize winners who were in the USA when they won their award than there are nobel prize winners who were outside the USA when they won their award. This supports BuzzFeed's claim that "most living Nobel laureates were based in the US when they won their prizes".     

### Exercise 4

```{r}
nobel_living_science <- nobel_living_science %>%
  mutate(
    born_country_us = if_else(born_country == "USA", "USA", "Other")
  )
```

```{r}
nobel_living_science %>%
  filter(born_country_us == "USA") %>%
  nrow()
```

105 winners were born in the US.

### Exercise 5

```{r}
ggplot(nobel_living_science, aes(x = country_us, fill = born_country_us)) +
  geom_bar() +
  facet_wrap(~ category) +
  coord_flip()
```

The data does not appear to support Buzzfeed's claim. Most of the US based Nobel laureates were born in the USA. 

### Exercise 6

```{r}
nobel_living_science %>%
  filter(country_us == "USA", born_country_us == "Other") %>%
  count(born_country) %>%
arrange(desc(n))
```

Germany and the United Kingdom are the most common.

