---
title: "Lab 03 - Nobel laureates"
author: "Shatavia Bellmon"
date: "1-31-26"
output: github_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
nobel <- read_csv("data/nobel.csv")
```

## Exercises

Here is a link to the [lab instructions](https://datascience4psych.github.io/DataScience4Psych/lab03.html).

### Exercise 1

view(nobel)

The nobel dataset contains `r nrow(nobel)` observations and `r ncol(nobel)` variables. Each row represents a a Nobel Prize winner and their information such as why they won the award and personal demographics.

### Exercise 2

```{r}

nobel_living <- nobel %>% 
  filter(
    is.na(died_date),
    !is.na(country),
    gender != "org"
  )
```

After filtering for living laureates with available country data and excluding organizations, the resulting data frame contains `r nrow(nobel_living)` observations.


### Exercise 3

Remove this text, and add your answer for Exercise 1 here.
Add code chunks as needed.
Don't forget to label your code chunk.
Do not use spaces in code chunk labels.

### Exercise 4

...

### Exercise 5

...

### Exercise 6

...
